\textbf{Advection versus diffusion in intraventricular transport}

### 2025-03-24 paper/main.tex
minor updates
abstract update

\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[2]{Inyoung Jeong}
\author[1,3]{Marie E.~Rognes}
\author[2, *]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway} % Kristian Augusts gate 23, 0164
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway} % Erling Skjalgsons gate 1, 7491
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}, \href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\date{\vspace{-2em}}

@@

    Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
    provides mechanical support of the brain and the flow of CSF helps
    distribute nutrients, neurotransmitters and metabolites throughout
    the central nervous system. CSF flow is partly driven
    by beating of motile cilia, hair-like structures that line the ependymal cell layer
    of the brain ventricles. Despite the physiological importance of CSF flow,
    the underlying mechanisms of cilia-mediated flow and transport
    in the brain ventricles remain yet to be comprehensively resolved.
    In this study, we analyze and evaluate the role of motile cilia
    in intraventricular CSF flow and transport. We use the geometry
    of the zebrafish larval brain ventricles, for which we have detailed knowledge of cilia properties and CSF motion velocities
    and frequencies, to develop finite element methods that model flow and transport.
    The computational model is validated by in vivo experiments that monitor transport
    of a photoconvertible protein secreted in the brain ventricles.
    Our results show that while cilia contribute to advection of large particles,
diffusion plays a significant role in the transport of small molecules such as proteins.
We also found that cilia location and the geometry of the ventricular system impact solute distribution.
Altogether, our work presents a computational framework that can be applied to other
ventricular systems, together with new concepts of how molecules are transported within the brain and its ventricles.

>>>>

\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[2]{Inyoung Jeong}
\author[1,3]{Marie E.~Rognes}
\author[2, *]{Nathalie Jurisch-Yaksi}
\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway} % Kristian Augusts gate 23, 0164
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway} % Erling Skjalgsons gate 1, 7491
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}, \href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\date{\vspace{-2em}}

@@

    Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
    provides mechanical support for the brain and the flow of CSF helps
    distribute nutrients, neurotransmitters and metabolites throughout
    the central nervous system. CSF flow is partly driven
    by beating of motile cilia, hair-like structures that line the ependymal cell layer
    of the walls of the brain ventricles. Despite the physiological importance of CSF flow,
    the underlying mechanisms of cilia-mediated flow and transport
    in the brain ventricles remain yet to be comprehensively resolved.
    In this study, we analyze and evaluate the role of motile cilia
    in intraventricular CSF flow and transport. We use the geometry
    of the zebrafish larval brain ventricles, for which we have detailed knowledge of cilia properties and CSF motion velocities
    and frequencies, to develop finite element methods that model flow and transport.
    The computational model is validated by in vivo experiments that monitor transport
    of a photoconvertible protein secreted in the brain ventricles.
    Our results show that while cilia contribute to advection of large particles,
diffusion plays a significant role in the transport of small molecules such as proteins.
We also found that cilia location and the geometry of the ventricular system impact solute distribution.
Altogether, our work presents a computational framework that can be applied to other
ventricular systems, together with new concepts of how molecules are transported within the brain and its ventricles.
